---
title: "project"
output: html_document
authors: Yishan Zheng, Rindala Fayyad, Lauren Mock, Daniel Herrera, Yuning Liu, Jinyi Che
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
STEP 1

Simulate data
- fix n 
- Xi from a jointly Uniform distribution
- Noise from a normal distribution (mean = 0, variance = very small)
- Yi linear combination of the Xi's + noise (with fixed Betas)

```{r}
simulation <- function(n, d) {
  set.seed(0)
  betas <- runif(d + 1, min = 0, max = 10)
  dat <- matrix(nrow = n, ncol = d + 1)
  dat <- as.data.frame(dat)
  e_vec <- rnorm(n, mean = 0, sd = 10)
  for (i in 1:n) {
    x_vec <- runif(d, min = 0, max = 100)
    y <- sum(betas[2 : length(betas)] * x_vec) + betas[1] + e_vec[i]
    row <- c(y, x_vec)
    dat[i,] <- row
  }
  colnames(dat)[1] <- "Y"
  lreg <- lm(Y ~ ., data = dat)
  return (list(estimate = summary(lreg)$coeff[,1], true = betas))
}

beta_estimate <- simulation(50, 1)

```


STEP 2

- Fit a linear regression model
- Get Beta estimators 

```{r}

```

Check Lab 10 simulation

REPEAT Step 1 and Step 2 100 times
Get expected value of the Beta estimators. 
Get Sample Variance of the Beta estimators. 
Get MSE (check for consistency)

Repeat again with larger p
Do this until you reach a problem

```{r}

```

